Tracing the Meta-Level: PyPy's Tracing JIT Compiler

### 2009-04-02 talk/icooolps2009/paper.tex
Add more references (slightly too large diff, as the reference-file is auto-generated).

The PyPy project\footnote{http://codespeak.net/pypy}
\cite{rigo_pypys_2006,carl_friedrich_bolz_to_2007} was started to implement a
new Python interpreter in Python but has now extended its goals to be an
environment where flexible implementation of dynamic languages can be written.
To implement a dynamic language with PyPy, an interpreter for that language has
to be written in RPython \cite{AACM-DLS07}. RPython ("Restricted Python") is a subset of Python
chosen in such a way that type inference can be performed on it. The language
interpreter can then be translated with the help of PyPy into various target
environments, such as C/Posix, the CLI and the JVM. This is done by a component
of PyPy called the \emph{translation toolchain}.

@@

Classical partial evaluation has failed to be useful for dynamic language for
much the same reasons why ahead-of-time compilers cannot compile them to
efficient code. If the partial evaluator knows only the program it simply does
not have enough information to produce good code. Therefore some work has been
done to do partial evaluation at runtime. One of the earliest works on runtime
specialisation is Tempo for C \cite{XXX}. However, it is essentially a normal
partial evaluator ``packaged as a library''; decisions about what can be
specialised and how are pre-determined. Another work in this direction is DyC
\cite{grant_dyc_2000}, another runtime specializer for C. Both of these projects
have a similar problem as DynamoRIO.  Targeting the C language makes
higher-level specialisation difficult (e.g.\ \texttt{malloc} can not be
optimized).

>>>>

The PyPy project\footnote{http://codespeak.net/pypy}
\cite{rigo_pypys_2006,carl_friedrich_bolz_to_2007} was started to implement a
new Python interpreter in Python but has now extended its goals to be an
environment where flexible implementation of dynamic languages can be written.
To implement a dynamic language with PyPy, an interpreter for that language has
to be written in RPython \cite{ancona_rpython:step_2007}. RPython ("Restricted Python") is a subset of Python
chosen in such a way that type inference can be performed on it. The language
interpreter can then be translated with the help of PyPy into various target
environments, such as C/Posix, the CLI and the JVM. This is done by a component
of PyPy called the \emph{translation toolchain}.

@@

Classical partial evaluation has failed to be useful for dynamic language for
much the same reasons why ahead-of-time compilers cannot compile them to
efficient code. If the partial evaluator knows only the program it simply does
not have enough information to produce good code. Therefore some work has been
done to do partial evaluation at runtime. One of the earliest works on runtime
specialisation is Tempo for C \cite{XXX}. However, it is essentially a normal
partial evaluator ``packaged as a library''; decisions about what can be
specialised and how are pre-determined. Another work in this direction is DyC
\cite{grant_dyc:expressive_2000}, another runtime specializer for C. Both of these projects
have a similar problem as DynamoRIO.  Targeting the C language makes
higher-level specialisation difficult (e.g.\ \texttt{malloc} can not be
optimized).
